---
title: "Reproducible Research Project 1"
author: "Raviraj"
date: "March 18, 2018"
output: html_document
---

```{r}
# 1.Code for reading in the dataset and/or processing the data
data<-read.csv("activity.csv")
```

Lets see the parameters & few rows of the data

```{r}
head(data)
```

It was asked to plot the no of steps taken each day, thus we will first group the given data based on date & summarize the no of steps taken each day

```{r}
# 2. Histogram of the total number of steps taken each day
library(dplyr)
data_steps <- data %>% filter(!is.na(steps)) %>% group_by(date) %>% summarize(steps = sum(steps))
head(data_steps)
```

## First Plot

```{r}
library(ggplot2)
ggplot(data_steps, aes(x = steps)) +
  geom_histogram(fill = "skyblue", binwidth = 1000) +
  labs(title = "Steps per Day Frequency", x = "Steps per day", y = "Frequency")
```

Lets see Mean & Median of the steps taken each day

```{r}
# 3. Mean and median number of steps taken each day
mean(data_steps$steps)
median(data_steps$steps)
```

Make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days.

```{r}
# 4. Time series plot of the average number of steps taken
data_interval<- data %>% select(steps, interval) %>% na.omit() %>% group_by(interval) %>% summarize(steps= mean(steps)) 
```

## Second Plot

```{r}
ggplot(data_interval, aes(x=interval, y=steps))+ geom_line(color = "firebrick")

```

The 5-minute interval that, on average, contains the maximum number of steps

```{r}
# 5. The 5-minute interval that, on average, contains the maximum number of steps
data_interval[which.max(data_interval$steps),]

```

Imputing missing data with mean
Lets see the total no of missing values
```{r}
missingVals <- sum(is.na(data))
```
Lets see the missing parameters
```{r}
names(which(sapply(data, anyNA)))
```

Replace the missed in column with the mean values

```{r}
# 6. Code to describe and show a strategy for imputing missing data
# imputing with mean values
replacewithmean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
meandata <- data%>% group_by(interval) %>% mutate(steps= replacewithmean(steps))
``` 
Check the new data after replacing with the mean values
```{r}
head(meandata)
sum(is.na(meandata))
```

## Third Plot

```{r}
# 7.Histogram of the total number of steps taken each day after missing values are imputed
steps_meandata<- meandata %>% group_by(date) %>% summarize(steps = sum(steps))
ggplot(steps_meandata, aes(x = steps))+geom_histogram(fill = "skyblue",bins = 20)+
        labs(title = "Steps per Day Frequency", x = "Steps per day", y = "Frequency")

```

```{r}
# 8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
meandata$day <- weekdays(as.Date(meandata$date))
meandata$type <- ifelse(meandata$day %in% c("Saturday","Sunday"),"weekend","weekday")
data_plot <- aggregate(meandata$steps, 
                       by=list(meandata$type, 
                               meandata$day, meandata$interval), mean)
colnames(data_plot)<-c("Type","Day","Interval","Steps")
library(lattice)
xyplot(Steps~Interval|Type, data=data_plot, type="l",  layout = c(1,2),
       main="Average Steps per Interval Based on Type of Day", 
       ylab="Average Number of Steps", xlab="Interval")
```












